---
title: "R Notebook"
output: html_notebook
---

This is my notebook for my final project. I will be making a web application using shiny to show campaign finance information for the Missouri Senate race.

#Load packages
```{r}
library(shiny)
library(rsconnect)
library(tidyverse)
library(janitor)
library(scales)
```

#Load the data
```{r}
contribs <- read_csv("data.1/contribs.csv")
```

#Documentation: 
https://www.fec.gov/data/elections/senate/MO/2022/

#make totals for each party
```{r}
party_contribs <- contribs %>% 
  filter(total_receipts!="0.00") %>%
  group_by(party_full) %>%
  summarise(total = sum(total_receipts))%>%
    mutate_each(funs(prettyNum(., big.mark=","))) %>% 
  mutate(party_full = str_to_title(party_full))
```

#totals for all candidates, then add commas
```{r}
names<- data.frame(names = c("Lucas Kunce", "Eric Schmitt", "Vicky J. Hartzler", "Dave Schatz", "Eric Greitens", "Mark T. Mccloskey", "Scott Sifton", "Spencer Ross Toder", "Timothy Jacob Shepard", "Jewel William Kelly Jr.", "Steven Craig Price", "Gena Ross", "Nicholas Crane Strauss", "Trudy Busch Valentine"))

all_cand_totals <- contribs %>% 
  filter(total_receipts!=0.00)%>%
  arrange(desc(total_receipts)) %>% 
  mutate(party_full = str_to_title(party_full)) %>% 
  mutate(names, .before=1)%>%  
  select(names, party_full, total_receipts)

#names<-sapply(strsplit(all_cand_totals$candidate_name, split=", "),function(x)  {paste(rev(x),collapse=" ")})  %>%
    #str_to_title())
#I got so frustrated by not knowing what to do with the Mr.s and Jr.s and the initials whatnot that I'm just going to write their names in a vector since there are only 14. 

all_cand_commas <- all_cand_totals %>%
    mutate_each(funs(prettyNum(., big.mark=",")))

all_cand_graph <- all_cand_totals %>% 
  select(names, total_receipts)
```


#totals for republican candidates
```{r}
rep_cand_totals <- all_cand_totals %>% 
  filter(party_full=="Republican Party")%>%  
  select(names, total_receipts)

rep_cand_commas <- rep_cand_totals %>% 
  mutate_each(funs(prettyNum(., big.mark=",")))
```
#totals for dems
```{r}

dem_cand_totals <- all_cand_totals %>% 
  filter(party_full=="Democratic Party")%>%  
  select(names, total_receipts)

dem_cand_commas <- dem_cand_totals %>% 
  mutate_each(funs(prettyNum(., big.mark=",")))
```
#totals for unk, independent
```{r}
unk_cand_totals <- all_cand_totals %>% 
  filter(party_full=="Independent" | party_full=="Unknown")%>%  
  select(names, total_receipts)

unk_cand_commas <- unk_cand_totals %>%
    mutate_each(funs(prettyNum(., big.mark=",")))
```


#Run the shiny app template
```{r}

ui <- fluidPage(
  
  h1("Missouri Senate Race"),
  
 verticalLayout(
   sidebarLayout(
    sidebarPanel( selectInput("var", 
                  label = "Choose a party to display",
                  choices = c("Republican", "Democrat",
                              "Unknown/Independent"),
                  selected = "Republican"),
                p("After Sen. Roy Blunt announced he would not be running for reelection for the U.S. Senate, the race for his seat quickly became crowded. Browse this application to learn more about the funding behind the race for Missouri's open Senate seat."),
p("This data is taken from the Federal Exchange Commission's website. It was accessed in early May."),
tags$a(href="https://www.fec.gov/data/elections/senate/MO/2022/", "Click here to access the FEC's campaign finance data.")
),
    
    mainPanel(
      tableOutput("table"),
      plotOutput("varPlot")
     )
  ),
),

 tags$hr(),  

h2("Totals"),
sidebarLayout(
    sidebarPanel(
      h4("Total contributions by party:"),
      tableOutput("TBL"),
      h4("Total contributions by candidate"),
      tableOutput("all_cands")
                     ),
    
    mainPanel(
      h3("Contribution sums for all candidates"),
      plotOutput("candPlot")
      )
  ),
tags$hr()  

)

server <- function(input, output){
  
  output$TBL <- renderTable({
      party_contribs
  })
  
  output$all_cands <- renderTable({
      all_cand_commas
  })
  
  output$table <- renderTable({data <- switch(input$var, 
                   "Republican" = rep_cand_commas,
                   "Democrat" = dem_cand_commas,
                   "Unknown/Independent" = unk_cand_commas)
  
})
  output$candPlot <- renderPlot(ggplot(all_cand_graph, aes(reorder(names, total_receipts))) +
  geom_bar(aes(weight = total_receipts), fill = "#3CAEA3") + 
  coord_flip() +
  ggtitle("Total candidate contributions") +
  xlab("Candidate") +
  ylab("Sum of contributions") +
  theme_bw(base_size = 16)+
    scale_y_continuous(labels = comma))

  
  output$varPlot <- renderPlot(ggplot({data <- switch(input$var,
                   "Republican" = rep_cand_totals,
                   "Democrat" = dem_cand_totals,
                   "Unknown/Independent" = unk_cand_totals)}, aes(reorder(names, total_receipts))) +
  geom_bar(aes(weight = total_receipts), fill = switch(input$var, 
                   "Republican" = "#ED5533",
                   "Democrat" = "#173F5F",
                   "Unknown/Independent" = "#F6D55C")) + 
  coord_flip() +
  ggtitle("Candidate contributions") +
  xlab("Candidate") +
  ylab("Sum of contributions") +
  theme_bw(base_size = 16)+
    scale_y_continuous(labels = comma))
  

}

shinyApp(ui=ui , server = server)

```

#Tasks!
- add reactive graphs!
- try to modify tables so they're easier to read -- column names, title case the cand names, add commas to the totals etc.
- add in a blurb about refunds, explaining Blunt negative
- style the web app (put non reactive total information in own column, make it prettier, etc.)
- add a normal p blurb about what the web app is, perhaps in the sidebar

#Note to Laura, tutorial timestamp 1:36:53 https://shiny.rstudio.com/tutorial/